Simultaneous Elicitation of Scoring Rule and Agent Preferences for Robust Winner Determination

### 2020-10-09 minimax.tex
small changes

More recently \citet{Llamazares2013} and \citet{Viappiani2018} %(among others \commentOC{Which others?}) 
considered the problem of dealing with unspecified weights in positional scoring rules; the first proposed a method based on Data Envelope Analysis, while the second proposed to aggregate the uncertainty in weights using criteria used in decision-making under uncertainty (such as minimax regret). 
The work of \citet{Cailloux2014} deals with the problem of eliciting a generic voting rule.

@@

Let ${\pprefeq_j}(x)$ designate the set of alternatives known to be considered by $j$ as less good than or equal to $x$, and ${\pprefinv_j}(y)$ be the set of alternatives known to be considered by $j$ as strictly better than $y$.
\begin{claim}
	The rank of $x$ in the PMR-maximizing linear orders of agent $j$ is $\hat{v}_{j}(x) = 1+\card{A}-\card{{\pprefeq_j}(x)}$, and the rank of $y$ is $\hat{v}_{j}(y)=1+\card{{\pprefinv_j}(y)}+\card{\beta}$, where $\card{\beta} = \card{A \setminus ({\pprefeq_j}(x) \cup {\pprefinv_j}(y))}$ if $(x \pprefeq_j y)$ and $\card{\beta} = 0$ otherwise.
\end{claim}
\begin{proof}
	The rank of $x$ is directly obtained from \cref{eq:complx}. The rank of $y$ is obtained by complementing \cref{eq:comply}, obtaining $a \prefeq_j y ⇔ (a \pprefeq_j y) ∨ ((x \pprefeq_j y) ∧ ¬(x \pprefeq_j a))$, and, observing that $a \pref_j y ⇔ a ≠ y ∧ a \prefeq_j y$, obtaining that $a \pref_j y$ if and only if
	\begin{equation}
		\label{eq:betteryinter}
		(a \neq y) ∧ [(a \pprefeq_j y) ∨ ((x \pprefeq_j y) ∧ ¬(x \pprefeq_j a))],
	\end{equation} 
	or equivalently, if and only if
	\begin{equation}
		\label{eq:bettery}
		%a \pref_j y ⇔ 
		(a \ppref_j y) ∨ ((x \pprefeq_j y) ∧ ¬(x \pprefeq_j a)).
	\end{equation} 
	Indeed, \eqref{eq:betteryinter} $⇒$ \eqref{eq:bettery}, and \eqref{eq:bettery} $⇒$ \eqref{eq:betteryinter} because $(x \pprefeq_j y) ∧ ¬(x \pprefeq_j a) ⇒ a ≠ y$ (as when $a = y$, $(x \pprefeq_j y)$ and $¬(x \pprefeq_j a)$ are opposite claims). Suffices now to rewrite \cref{eq:bettery} to let the two disjuncts designate disjoint sets:
	\begin{equation}
		\label{eq:betteryfinal}
		a \pref_j y ⇔ 
		(a \ppref_j y) ∨ ((x \pprefeq_j y) ∧ ¬(x \pprefeq_j a) ∧ ¬(a \ppref_j y)).
	\end{equation}
\end{proof}

The claim can also be understood by observing that in the case $(x \pprefeq_j y)$, $\beta$ is the number of alternatives incomparable with both $x$ and $y$.

\begin{claim}
	The $\PMR$ can be written as:
	\begin{align} 
		\PMR(x,y; \pprofile, W)  
		& = \max_{w \in W} \sum_{j=1}^n w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} = \\ 
		& = \max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i. 
	\end{align}
	where $\hat{\alpha}_{r}^{y}$ (resp. $\hat{\alpha}_{r}^{x}$)  is the number of times $y$ (resp. $x$) is at rank $r$ in the complete profile $\hat{\profile}$.
\end{claim}
The last claim shows that PMR is linear in the weights.
Recall that the preferences of the chair are encoded with linear constraints $\Co_{W}$.
The pairwise max regret $\PMR(x,y; \pprofile,W)$ can be obtained as the solution of the following linear program defined on the variables $w_1, …, w_m$, which represent the weights attached to different positions:

\begin{align}
	\max_{\w} & \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r}\\
	\text{ s.t. } & \text{\cref{eq:monotone}} \text{ and } \text{\cref{eq:convexity}} \text{ and } \Co_W.
\end{align}
Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables 
(we leave $w_{1}$ and $w_{m}$ in the LP just for clarity of presentation).

>>>>

More recently \citet{Llamazares2013} and \citet{Viappiani2018} %(among others \commentOC{Which others?}) 
considered the problem of dealing with unspecified weights in positional scoring rules; the first proposed a method based on Data Envelopment Analysis (DEA), while the second proposed to aggregate the uncertainty in weights using criteria used in decision-making under uncertainty (such as minimax regret). 
The work of \citet{Cailloux2014} deals with the problem of eliciting a generic voting rule.

@@

Let ${\pprefeq_j}(x)$ designate the set of alternatives known to be considered by $j$ as less good than or equal to $x$, and ${\pprefinv_j}(y)$ be the set of alternatives known to be considered by $j$ as strictly better than $y$.
\begin{claim}
	The rank of $x$ in the PMR-maximizing linear orders of agent $j$ is $\hat{v}_{j}(x) = 1+\card{A}-\card{{\pprefeq_j}(x)}$, and the rank of $y$ is $\hat{v}_{j}(y)=1+\card{{\pprefinv_j}(y)}+\card{\beta}$, where $\card{\beta} = \card{A \setminus ({\pprefeq_j}(x) \cup {\pprefinv_j}(y))}$ if $(x \pprefeq_j y)$ and $\card{\beta} = 0$ otherwise.
	\label{claim:rankPMR}
\end{claim}
\begin{proof}
	The rank of $x$ is directly obtained from \cref{eq:complx}. The rank of $y$ is obtained by complementing \cref{eq:comply}, obtaining $a \prefeq_j y ⇔ (a \pprefeq_j y) ∨ ((x \pprefeq_j y) ∧ ¬(x \pprefeq_j a))$, and, observing that $a \pref_j y ⇔ a ≠ y ∧ a \prefeq_j y$, obtaining that $a \pref_j y$ if and only if
	\begin{equation}
		\label{eq:betteryinter}
		(a \neq y) ∧ [(a \pprefeq_j y) ∨ ((x \pprefeq_j y) ∧ ¬(x \pprefeq_j a))],
	\end{equation} 
	or equivalently, if and only if
	\begin{equation}
		\label{eq:bettery}
		%a \pref_j y ⇔ 
		(a \ppref_j y) ∨ ((x \pprefeq_j y) ∧ ¬(x \pprefeq_j a)).
	\end{equation} 
	Indeed, \eqref{eq:betteryinter} $⇒$ \eqref{eq:bettery}, and \eqref{eq:bettery} $⇒$ \eqref{eq:betteryinter} because $(x \pprefeq_j y) ∧ ¬(x \pprefeq_j a) ⇒ a ≠ y$ (as when $a = y$, $(x \pprefeq_j y)$ and $¬(x \pprefeq_j a)$ are opposite claims). Suffices now to rewrite \cref{eq:bettery} to let the two disjuncts designate disjoint sets:
	\begin{equation}
		\label{eq:betteryfinal}
		a \pref_j y ⇔ 
		(a \ppref_j y) ∨ ((x \pprefeq_j y) ∧ ¬(x \pprefeq_j a) ∧ ¬(a \ppref_j y)).
	\end{equation}
\end{proof}
The claim can also be understood by observing that in the case $(x \pprefeq_j y)$, $\beta$ is the number of alternatives incomparable with both $x$ and $y$.
\begin{claim}
	The $\PMR$ can be written as:
	\begin{align} 
		\PMR(x,y; \pprofile, W)  
		& = \max_{w \in W} \sum_{j=1}^n w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} = \\ 
		& = \max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i. 
	\end{align}
	where $\hat{\alpha}_{r}^{y}$ (resp. $\hat{\alpha}_{r}^{x}$)  is the number of times $y$ (resp. $x$) is at rank $r$ in the complete profile $\hat{\profile}$ as determined in Claim \ref{claim:rankPMR}. %\ref{claim:completion}.
\end{claim}
The last claim shows that PMR is linear in the weights.
Recall that the preferences of the chair are encoded with linear constraints $\Co_{W}$.
The pairwise max regret $\PMR(x,y; \pprofile,W)$ can be obtained as the solution of the following linear program defined on the variables $w_1, …, w_m$, which represent the weights attached to different positions:
\begin{align}
	\max_{\w} & \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r}\\
	\text{ s.t. } & \text{\cref{eq:monotone}} \text{ and } \text{\cref{eq:convexity}} \text{ and } \Co_W.
\end{align}
Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables 
(we leave $w_{1}$ and $w_{m}$ in the LP just for clarity of presentation).
